### 2022-03-24 draft.tex
Deploying to gh-pages from @ emuise/ms-assessingRepresentation@aab9ee912f5afa1c088b42fec9a648ca01ed9f79 🚀

The province of British Columbia, Canada, covers 94.4 million ha, of
which approximately 64\% is forested (BC Ministry of Forests 2003), and
encapsulates a wide variety of biomes and ecosystems. This diversity of
ecosystems is in part due to the large area as well as variations in
topography and climate. The existing Biogeoclimatic Ecosystem
Classification (BEC) system disaggregates BC ecosystems into zones
(Pojar et al. 1987). The broadest classification delineates 16 zones,
which are further broken down into subzones, variants, and phases based
on microclimate, precipitation, and topography (Pojar et al. 1987,
Meidinger and Pojar 1991). As a result, BEC zones vary widely in size
(ranging from 0.25 million ha to 17.5 million ha), and in number of
subzones (from 1 to 43; see Table 1).

@@

Forest structural attributes vary between PA and UA in BC
(fig.~\ref{fig:structure-3d-scatter}). The largest differences between
PA and UA are found in canopy structure in the Coastal Douglas-fir BEC
zone, with the protected area having much higher canopy structure
values. As shown in fig.~\ref{fig:t-test-plot}, forests are commonly
significantly different when comparing PA vs UA across all attributes.
When examining the forests on an BEC zone level, only one BEC zone has a
\textgreater5\% difference in vertical forest structure (co-efficient of
variation in vegetation returns), six BEC zones have \textgreater5\%
difference in canopy cover, and five BEC zones have a \textgreater5\%
difference in canopy height. Ponderosa pine has large differences in
canopy cover and canopy height (\textgreater5\%), but minor differences
in elevation covariance (only 0.25\%; Table 2). PA in the Ponderosa
Pine, Interior Mountain Heather Alpine, and Coastal Douglas-fir have
more aboveground biomass than in UA in corresponding areas
(fig.~\ref{fig:structure-3d-scatter}).

@@

In high elevation ecosystems, Boreal Altai Fescue Alpine dominates the
PA proportions above 3000m, replacing the Coastal Mountain-Heather
Alpine ecosystem found in UA (fig.~\ref{fig:bec-elev}). These zones were
both protected at rates above the average (fig.~\ref{fig:bec-conch}),
and above the Aichi biodiversity targets. Interior mountain-heather
alpine had large differences in canopy cover and canopy height, while
Boreal Altai Fescue Alpine only showed large differences in height. The
Coastal Mountain-heather Alpine did not any have large forest structural
attribute differences (Table 2).

@@

Utilizing this information on the proportion of BEC zones protected
(fig.~\ref{fig:bec-conch}), as well as their forest structural
attributes (Table 2), it is possible to identify which forest structures
need to be added to the PA network in BC. Those BEC zones with large
differences (identified as being \textgreater5\% change from PA to UA)
suggest additional protection is needed to encapsulate these
underrepresented forest structures. For example: the forests in the
Bunchgrass zone have large differences in both canopy cover and canopy
height, with the PA having larger values in both attributes (Table 2).
New PA in this BEC zone should contain forests with shorter and more
open forests. A future avenue of research could be to incorporate forest
structural attributes into spatially optimized PA placement approaches
(Christensen et al. 2009).

>>>>

The province of British Columbia, Canada, covers 94.4 million ha, of
which approximately 64\% is forested (BC Ministry of Forests 2003), and
encapsulates a wide variety of biomes and ecosystems. This diversity of
ecosystems is in part due to the large area as well as variations in
topography and climate. The existing Biogeoclimatic Ecosystem
Classification (BEC) system disaggregates BC ecosystems into zones
(Pojar et al. 1987). The broadest classification delineates 16 zones,
which are further broken down into subzones, variants, and phases based
on microclimate, precipitation, and topography (Pojar et al. 1987,
Meidinger and Pojar 1991). As a result, BEC zones vary widely in size
(ranging from 0.25 million ha to 17.5 million ha), and in number of
subzones (from 1 to 43; see (\textbf{tab:bec-table?})).

\hypertarget{tbl:bec-table}{}
\begin{longtable}[]{@{}lllll@{}}
\caption{\label{tbl:bec-table}Number of subzones, total area, and
percent protected by BEC Zone.}\tabularnewline
\toprule
Zone & Zone Name & \# of Subzones & Area (ha) & \%
Protected\tabularnewline
\midrule
\endfirsthead
\toprule
Zone & Zone Name & \# of Subzones & Area (ha) & \%
Protected\tabularnewline
\midrule
\endhead
BAFA & Boreal Altai Fescue Alpine & 2 & 6,286,778 &
30.1\%\tabularnewline
BG & Bunchgrass & 2 & 257,072 & 11.8\%\tabularnewline
BWBS & Boreal White and Black Spruce & 5 & 16,404,142 &
8.6\%\tabularnewline
CDF & Coastal Douglas-fir & 1 & 251,232 & 4.8\%\tabularnewline
CMA & Coastal Mountain-heather Alpine & 3 & 3,574,039 &
17.9\%\tabularnewline
CWH & Coastal Western Hemlock & 10 & 10,795,067 & 19.5\%\tabularnewline
ESSF & Engelmann Spruce -- Subalpine Fir & 43 & 17,465,113 &
17.8\%\tabularnewline
ICH & Interior Cedar -- Hemlock & 12 & 5,538,842 & 10.2\%\tabularnewline
IDF & Interior Douglas-fir & 12 & 4,488,085 & 5.9\%\tabularnewline
IMA & Interior Mountain-heather Alpine & 2 & 1,257,949 &
29.2\%\tabularnewline
MH & Mountain Hemlock & 6 & 4,059,301 & 19.8\%\tabularnewline
MS & Montane Spruce & 8 & 2,863,394 & 9.4\%\tabularnewline
PP & Ponderosa Pine & 1 & 294,985 & 7.1\%\tabularnewline
SBPS & Sub-Boreal Pine -- Spruce & 4 & 2,265,365 & 9.5\%\tabularnewline
SBS & Sub-Boreal Spruce & 11 & 10,337,497 & 6.7\%\tabularnewline
SWB & Spruce -- Willow -- Birch & 6 & 8,655,855 & 23.3\%\tabularnewline
\bottomrule
\end{longtable}

@@

Forest structural attributes vary between PA and UA in BC
(fig.~\ref{fig:structure-3d-scatter}). The largest differences between
PA and UA are found in canopy structure in the Coastal Douglas-fir BEC
zone, with the protected area having much higher canopy structure
values. As shown in fig.~\ref{fig:t-test-plot}, forests are commonly
significantly different when comparing PA vs UA across all attributes.
When examining the forests on an BEC zone level, only one BEC zone has a
\textgreater5\% difference in vertical forest structure (co-efficient of
variation in vegetation returns), six BEC zones have \textgreater5\%
difference in canopy cover, and five BEC zones have a \textgreater5\%
difference in canopy height. Ponderosa pine has large differences in
canopy cover and canopy height (\textgreater5\%), but minor differences
in elevation covariance (only 0.25\%; tbl.~\ref{tbl:vector-table}). PA
in the Ponderosa Pine, Interior Mountain Heather Alpine, and Coastal
Douglas-fir have more aboveground biomass than in UA in corresponding
areas (fig.~\ref{fig:structure-3d-scatter}).

Zone

Elevation Covariance

Canopy Cover (\%)

Canopy Height (m)

PA

UA

\% Difference

PA

UA

\% Difference

PA

UA

\% Difference

BAFA

0.39

0.39

0.01\%

46.94\%

48.47\%

3.17\%

14.03

13.11

-7\%

BG

0.38

0.38

-1.41\%

61.8\%

58.71\%

-5.26\%

23.00

21.58

-6.59\%

BWBS

0.37

0.38

2.62\%

67.18\%

64.72\%

-3.8\%

15.83

15.69

-0.9\%

CDF

0.31

0.33

6.35\%

89.38\%

83.69\%

-6.79\%

27.35

26.58

-2.89\%

CMA

0.38

0.39

1.63\%

62.65\%

64.01\%

2.13\%

19.91

20.01

0.48\%

CWH

0.34

0.33

-3.83\%

83.94\%

85.26\%

1.55\%

21.58

22.34

3.4\%

ESSF

0.37

0.37

0.34\%

61.7\%

64.97\%

5.04\%

18.90

18.91

0.07\%

ICH

0.36

0.36

-1.75\%

81.24\%

83.52\%

2.73\%

22.39

22.18

-0.98\%

IDF

0.36

0.36

-0.3\%

67.42\%

67.9\%

0.71\%

21.98

22.49

2.29\%

IMA

0.38

0.36

-3.72\%

68.17\%

62.07\%

-9.83\%

22.53

21.06

-6.98\%

MH

0.36

0.36

0.25\%

76.87\%

77.85\%

1.26\%

19.42

18.31

-6.07\%

MS

0.35

0.35

0.31\%

57.99\%

60.41\%

4.01\%

20.64

20.86

1.04\%

PP

0.36

0.37

0.25\%

57.92\%

48.93\%

-18.36\%

19.88

18.03

-10.24\%

SBPS

0.36

0.35

-1.97\%

32.98\%

34.63\%

4.76\%

18.00

18.70

3.75\%

SBS

0.37

0.37

-0.4\%

62.24\%

67.25\%

7.45\%

18.51

18.67

0.82\%

SWB

0.39

0.39

-1.22\%

56.67\%

57.71\%

1.8\%

13.78

13.67

-0.83\%

: Mean values of forest structural attributes in protected areas (PA),
unprotected areas (UA), as well as the percent difference between the
means. Zones with more than a 5\% difference are bolded.
\{\#tbl:vector-table\}

@@

In high elevation ecosystems, Boreal Altai Fescue Alpine dominates the
PA proportions above 3000m, replacing the Coastal Mountain-Heather
Alpine ecosystem found in UA (fig.~\ref{fig:bec-elev}). These zones were
both protected at rates above the average (fig.~\ref{fig:bec-conch}),
and above the Aichi biodiversity targets. Interior mountain-heather
alpine had large differences in canopy cover and canopy height, while
Boreal Altai Fescue Alpine only showed large differences in height. The
Coastal Mountain-heather Alpine did not any have large forest structural
attribute differences (tbl.~\ref{tbl:vector-table}).

@@

Utilizing this information on the proportion of BEC zones protected
(fig.~\ref{fig:bec-conch}), as well as their forest structural
attributes (tbl.~\ref{tbl:vector-table}), it is possible to identify
which forest structures need to be added to the PA network in BC. Those
BEC zones with large differences (identified as being \textgreater5\%
change from PA to UA) suggest additional protection is needed to
encapsulate these underrepresented forest structures. For example: the
forests in the Bunchgrass zone have large differences in both canopy
cover and canopy height, with the PA having larger values in both
attributes (tbl.~\ref{tbl:vector-table}). New PA in this BEC zone should
contain forests with shorter and more open forests. A future avenue of
research could be to incorporate forest structural attributes into
spatially optimized PA placement approaches (Christensen et al. 2009).
